The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-03-17 paper/iphas-dr2.tex
Assign subfigures to Fig. 10
The two plots in Fig. 10 were referenced separately in the text, but were appearing as Fig. 6.2 (the section containing the reference). The subcaption package has been added to give a/b subfigure lettering, and the caption package to set the letters not to display. \captionsetup can be commented out to have the letters appear.

This does have the unfortunate effect of (slightly) increasing the space between the figures and the caption.

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{amssymb,amsmath}
\usepackage[pdftex]{graphicx}
\usepackage{longtable} 
\usepackage{float}
\usepackage{booktabs}
\usepackage{aas_macros}

@@

\begin{abstract} % BACKGROUND, OBJECTIVE, METHODS, RESULTS, CONCLUSIONS
The INT/WFC Photometric H$\alpha$ Survey 
of the Northern Galactic Plane (IPHAS)
is a 1800 deg$^2$ imaging survey
covering the entire northern Milky Way at $|b| < 5^\circ$
in the $r$, $i$ and \ha\ filters 
using the Wide Field Camera (WFC) 
on the 2.5-meter Isaac Newton Telescope (INT) in La Palma.
We present the first quality-controlled and
uniformly-calibrated source catalogue
to have been extracted from the images,
providing single-epoch photometry
for 219 million unique sources
across 92\% of the survey area.
The observations were carried out between 2003 and 2013
at a median seeing of 1.1 arcsec
and to a mean depth of $r=21.2$ / $i=20.0$ / $\ha=20.3$
($5\sigma$ limits, Vega system).
We explain the data reduction 
and quality control procedures,
describe and test the new uniform calibration,
and detail the construction of the 97-column source catalogue.
We show that the global calibration is accurate to
0.03~mag (rms) by validating against SDSS,
and provide a series of recommended quality criteria
which can be used to select the most reliable data
from the catalogue.
Finally, we demonstrate the ability of the 
catalogue's unique
$(r-i)$/($r-\ha)$ colour-colour diagram to
(i) characterise stellar populations and extinction regimes
towards different Galactic sight-lines
and (ii) select candidate \ha\ emission-line objects.
IPHAS is the first survey to offer comprehensive CCD photometry
of point sources across the Galactic Plane at visible wavelengths,
providing the much-needed counterpart
to infrared surveys.
We also provide images for some of the most
crowded regions to be faced by Gaia.
\end{abstract}

@@

The detailed properties of the IPHAS observing programme 
have been presented before 
by \citet{Drew2005} and \citet{Gonzalez-Solares2008}. 
To set the stage for this release, we briefly remind of some key points.
IPHAS is a 1800~sq. deg. imaging survey of the northern Galactic Plane, 
providing images and photometry in Sloan $r, i$ 
along with narrowband H$\alpha$. 
It is carried out using the Wide Field Camera (WFC) 
on the 2.5-meter Isaac Newton Telescope (INT) in La Palma. 
It is the first digital survey to offer comprehensive CCD photometry
of point sources in the Galactic Plane at visible wavelengths, 
and does so down to a limiting magnitude of $\sim$20th.
The IPHAS footprint on the northern sky spans a box 
of roughly 180 by 10 degrees, 
taking in the entire northern Galactic Plane 
at latitudes $-5^{\circ} < b < +5^{\circ}$ 
and longitudes $30^{\circ} < l < 215^{\circ}$.

@@

\begin{figure*}
    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_rshift.pdf} 
    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_r.pdf} 
    \caption{Median magnitude offset in the $r$ band between IPHAS and APASS,
             plotted on a field-by-field basis
             prior to the re-calibration procedure.
             Each square represents the footprint of an IPHAS field
             which contains at least 30 stars with a counterpart
             in the APASS DR7 catalogue.
             The colours denote the median
             IPHAS-APASS magnitude offset in each field,
             which was computed after applying the APASS-to-IPHAS
             transformation to the APASS magnitudes (Eqn.~\ref{eqn:apass_r}).
             For clarity, we do not show the fields at the offset positions.}
        \label{fig:apass_r}
    \vspace{1cm}
    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_rshift.pdf}
    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_r.pdf} 
    \caption{Median magnitude offset in the $r$ band
             between IPHAS and SDSS after the re-calibration
             procedure was applied.
             Each square represents the footprint of an IPHAS field
             which contains at least 30 stars
             with a counterpart in the SDSS catalogue.
             The colours denote the median IPHAS-SDSS magnitude offset
             in each field,
             which was computed after applying the SDSS-to-IPHAS
             transformation to the SDSS magnitudes (Eqn.~\ref{eqn:sdss_r}).
             For clarity, we do not show the fields at the offset positions.}
    \label{fig:sdss_r}
\end{figure*}

@@

As explained above, we enforce a strong external constraint
on the solution by keeping the zeropoint fixed 
for the 6\,756 fields which have been compared
and calibrated against APASS earlier.
We hereafter refer to these fields as \emph{anchors}.
It is asserted that the zeropoints $a_i$ of these anchor fields 
are known and not solved for,
though they do appear in the vector $b_j$ as constraints.
In addition to the APASS-based anchors, 
we selected 3\,273 additional anchor fields by hand
to provide additional constraints in regions not covered by APASS.
These extra anchors were deemed to have reliable zeropoints 
based on 
(i) the information contained in the observing logs,
(ii) the stability of the standard star zeropoints during the night, and
(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
which is located at $\sim$500\,meter from the INT.

@@

Information on the quality of each detection is included 
in a series of additional columns.
We draw attention to three useful flags
which warn about the likely presence of a systematic error:
\begin{enumerate}
\item The \emph{saturated} column is used to flag sources
for which the peak pixel height exceeds 55000 counts,
which is typically the case for stars brighter than 12-13th magnitude.
Although the pipeline attempts to extrapolate the brightness of
saturated stars based on the shape of the PSF,
such extrapolation is prone to error
and we do not recommend its use.
\item The \emph{deblend} column is used to flag sources 
which partially overlap with a nearby neighbour.
Although the pipeline applies a deblending procedure
to such objects, the procedure is currently applied separately
in each band and hence the $(r-i)$ and $(r-\ha)$ colours
may be inaccurate if the deblending proceeded differently in each band.
\item The \emph{brightNeighb} column is used to flag sources which are located
within 5 arcmin from an object brighter than $V=7$ 
according to the Bright Star Catalogue (BSC; Hoffleit et al. 1991), 
or within 10 arcmin if the neighbour is brighter than $V=4$.
Such very bright stars are known to cause systematic errors
and spurious detections as a result of stray light 
and diffraction spikes.
\end{enumerate}
In addition to the above, we also created warning flags for internal bookkeeping.
For example, we flagged detections which fell in the strongly vignetted regions of the focal plane,
which were truncated by CCD edges,
or which were otherwise affected by bad pixels in the detector.
We will explain below that none of such detections 
have been included in the catalogue.
An alternative detection was available in essentially all these situations
as a result of the IPHAS field pair strategy,
and hence it was not necessary to include these internal warning flags
in the source catalogue.

@@

   pStar $>$ 0.9 &
   145 million (66\%) &
   % pStar > 0.89 is identical to mergedClass == -1
   The object must appears as a perfect point source,
   as inferred from comparing its Point Spread Function (PSF)
   with the average PSF measured in the same CCD. \\

@@

\begin{figure*}
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=\textwidth]{./plots/pairmag/pairmag.pdf}
\label{fig:pairmag}
\end{minipage}
\hspace{0.5cm}
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=\textwidth]{./plots/pairmag/pairmag-reliable.pdf}
\label{fig:pairmag_reliable}
\end{minipage}
\caption{Photometric repeatability as a function of magnitude
         for both all sources in the catalogue (left panel)
         and for the \emph{veryReliable} sources alone (right panel).
         The values shown are the mean residuals
         between the primary and the secondary detections,
         while the error-bars show the standard deviations.
         Left panel: 
         the best photometric repeatability is reached at $r=14$
         with $\sigma_{r-r2}~=~0.041$~mag.
         Bright stars at $r<13$ and $i<12$ 
         show increasing uncertainties due to saturation effects.
         Right panel: we find that applying the quality criteria
         has improved the photometric repeatability significantly.
         The best repeatability is again reached at $r=14$
         but has reduced to $\sigma=0.028$ mag.
         The quality criteria have also been successful
         at removing objects with large systematics at the bright
         and faint ends.
    }
\end{figure*}

>>>>

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{amssymb,amsmath}
\usepackage[pdftex]{graphicx}
\usepackage{longtable} 
\usepackage{float}
\usepackage{booktabs}
\usepackage{aas_macros}
\usepackage{caption}
\usepackage{subcaption}

@@

\begin{abstract} % BACKGROUND, OBJECTIVE, METHODS, RESULTS, CONCLUSIONS
The INT/WFC Photometric H$\alpha$ Survey 
of the Northern Galactic Plane (IPHAS)
is a 1800 deg$^2$ imaging survey
covering the entire northern Milky Way at $|b| < 5^\circ$
in the $r$, $i$ and \ha\ filters 
using the Wide Field Camera (WFC) 
on the 2.5-metre Isaac Newton Telescope (INT) in La Palma.
We present the first quality-controlled and
uniformly-calibrated source catalogue
to have been extracted from the images,
providing single-epoch photometry
for 219 million unique sources
across 92\% of the survey area.
The observations were carried out between 2003 and 2013
at a median seeing of 1.1 arcsec
and to a mean depth of $r=21.2$ / $i=20.0$ / $\ha=20.3$
($5\sigma$ limits, Vega system).
We explain the data reduction 
and quality control procedures,
describe and test the new uniform calibration,
and detail the construction of the 97-column source catalogue.
We show that the global calibration is accurate to
0.03~mag (rms) by validating against SDSS,
and provide a series of recommended quality criteria
which can be used to select the most reliable data
from the catalogue.
Finally, we demonstrate the ability of the 
catalogue's unique
$(r-i)$/($r-\ha)$ colour-colour diagram to
(i) characterise stellar populations and extinction regimes
towards different Galactic sight-lines
and (ii) select candidate \ha\ emission-line objects.
IPHAS is the first survey to offer comprehensive CCD photometry
of point sources across the Galactic Plane at visible wavelengths,
providing the much-needed counterpart
to infrared surveys.
We also provide images for some of the most
crowded regions to be faced by Gaia.
\end{abstract}

@@

The detailed properties of the IPHAS observing programme 
have been presented before 
by \citet{Drew2005} and \citet{Gonzalez-Solares2008}. 
To set the stage for this release, we briefly remind of some key points.
IPHAS is a 1800~sq. deg. imaging survey of the northern Galactic Plane, 
providing images and photometry in Sloan $r, i$ 
along with narrowband H$\alpha$. 
It is carried out using the Wide Field Camera (WFC) 
on the 2.5-metre Isaac Newton Telescope (INT) in La Palma. 
It is the first digital survey to offer comprehensive CCD photometry
of point sources in the Galactic Plane at visible wavelengths, 
and does so down to a limiting magnitude of $\sim$20th.
The IPHAS footprint on the northern sky spans a box 
of roughly 180 by 10 degrees, 
taking in the entire northern Galactic Plane 
at latitudes $-5^{\circ} < b < +5^{\circ}$ 
and longitudes $30^{\circ} < l < 215^{\circ}$.

@@

\begin{figure*}
    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_rshift.pdf} 
    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_r.pdf} 
    \caption{Median magnitude offset in the $r$ band between IPHAS and APASS,
             plotted on a field-by-field basis
             prior to the re-calibration procedure.
             Each square represents the footprint of an IPHAS field
             which contains at least 30 stars with a counterpart
             in the APASS DR7 catalogue.
             The colours denote the median
             IPHAS-APASS magnitude offset in each field,
             which was computed after applying the APASS-to-IPHAS
             transformation to the APASS magnitudes (Eqn.~\ref{eqn:apass_r}).
             For clarity, we do not show the fields at the offset positions.}
        \label{fig:apass_r}
    \vspace{1cm}
    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_rshift.pdf}
    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_r.pdf} 
    \caption{Median magnitude offset in the $r$ band
             between IPHAS and SDSS after the re-calibration
             procedure was applied.
             Each square represents the footprint of an IPHAS field
             which contains at least 30 stars
             with a counterpart in the SDSS DR9 catalogue.
             The colours denote the median IPHAS-SDSS magnitude offset
             in each field,
             which was computed after applying the SDSS-to-IPHAS
             transformation to the SDSS magnitudes (Eqn.~\ref{eqn:sdss_r}).
             For clarity, we do not show the fields at the offset positions.}
    \label{fig:sdss_r}
\end{figure*}

@@

As explained above, we enforce a strong external constraint
on the solution by keeping the zeropoint fixed 
for the 6\,756 fields which have been compared
and calibrated against APASS earlier.
We hereafter refer to these fields as \emph{anchors}.
It is asserted that the zeropoints $a_i$ of these anchor fields 
are known and not solved for,
though they do appear in the vector $b_j$ as constraints.
In addition to the APASS-based anchors, 
we selected 3\,273 additional anchor fields by hand
to provide additional constraints in regions not covered by APASS.
These extra anchors were deemed to have reliable zeropoints 
based on 
(i) the information contained in the observing logs,
(ii) the stability of the standard star zeropoints during the night, and
(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
which is located $\sim$500\,metres from the INT.

@@

Information on the quality of each detection is included 
in a series of additional columns.
We draw attention to three useful flags
which warn about the likely presence of a systematic error:
\begin{enumerate}
\item The \emph{saturated} column is used to flag sources
for which the peak pixel height exceeds 55000 counts,
which is typically the case for stars brighter than 12-13th magnitude.
Although the pipeline attempts to extrapolate the brightness of
saturated stars based on the shape of the PSF,
such extrapolation is prone to error
and we do not recommend its use.
\item The \emph{deblend} column is used to flag sources 
which partially overlap with a nearby neighbour.
Although the pipeline applies a deblending procedure
to such objects, the procedure is currently applied separately
in each band and hence the $(r-i)$ and $(r-\ha)$ colours
may be inaccurate if the deblending proceeded differently in each band.
\item The \emph{brightNeighb} column is used to flag sources which are located
within 5 arcmin from an object brighter than $V=7$ 
according to the Bright Star Catalogue (BSC; Hoffleit et al. 1991), 
or within 10 arcmin if the neighbour is brighter than $V=4$.
Such very bright stars are known to cause systematic errors
and spurious detections as a result of stray light 
and diffraction spikes.
\end{enumerate}
In addition to the above, we also created warning flags for internal bookkeeping.
For example, we flagged detections which fell in the strongly vignetted regions of the focal plane,
which were truncated by CCD edges,
or which were otherwise affected by bad pixels in the detector.
We will explain below that no such detections 
have been included in the catalogue.
An alternative detection was available in essentially all these situations
as a result of the IPHAS field pair strategy,
and hence it was not necessary to include these internal warning flags
in the source catalogue.

@@

   pStar $>$ 0.9 &
   145 million (66\%) &
   % pStar > 0.89 is identical to mergedClass == -1
   The object must appear as a perfect point source,
   as inferred from comparing its Point Spread Function (PSF)
   with the average PSF measured in the same CCD. \\

@@

\begin{figure*}
\captionsetup[subfigure]{labelformat=empty}
\begin{subfigure}[b]{0.45\linewidth}
\centering
\includegraphics[width=\textwidth]{./plots/pairmag/pairmag.pdf}
\caption{}
\label{fig:pairmag}
\end{subfigure}
\hspace{0.5cm}
\begin{subfigure}[b]{0.45\linewidth}
\centering
\includegraphics[width=\textwidth]{./plots/pairmag/pairmag-reliable.pdf}
\caption{}
\label{fig:pairmag_reliable}
\end{subfigure}
\caption{Photometric repeatability as a function of magnitude
         for both all sources in the catalogue (left panel)
         and for the \emph{veryReliable} sources alone (right panel).
         The values shown are the mean residuals
         between the primary and the secondary detections,
         while the error-bars show the standard deviations.
         Left panel: 
         the best photometric repeatability is reached at $r=14$
         with $\sigma_{r-r2}~=~0.041$~mag.
         Bright stars at $r<13$ and $i<12$ 
         show increasing uncertainties due to saturation effects.
         Right panel: we find that applying the quality criteria
         has improved the photometric repeatability significantly.
         The best repeatability is again reached at $r=14$
         but has reduced to $\sigma=0.028$ mag.
         The quality criteria have also been successful
         at removing objects with large systematics at the bright
         and faint ends.
    }
\end{figure*}
